---
title: "Red Wine Quality Analysis"
author: "Author : Animesh Chowdhury"
output:
  html_document: 
    toc: true # Creates table of contents
    theme: united
    toc_depth: 3
    toc_float: true
---

**Dataset Overview: **
This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine.The inputs include objective tests (e.g. PH values) and the output is based on sensory data (median of at least 3 evaluations made by wine experts).At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

**Variables: **
Input variables (based on physicochemical tests):
fixed acidity (tartaric acid - g / dm^3),
volatile acidity (acetic acid - g / dm^3),
citric acid (g / dm^3),
residual sugar (g / dm^3),
chlorides (sodium chloride - (g / dm^3),
free sulfur dioxide (mg / dm^3),
total sulfur dioxide (mg / dm^3),
density (g / cm^3),
pH,
sulphates (potassium sulphate - g / dm3),
alcohol (% by volume),
   
Output variable (based on sensory data): 
quality (score between 0 and 10)

**Scope of Analysis: **
In this analysis we will try to evaluate the chemical properties which affect the quality of red wine.

```{r setup, include=FALSE}
# To hide the code in the knitted html output
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r echo = FALSE,  message = FALSE, warning = FALSE}
#Setting path for the root folder
setwd("/Users/animesh01/Desktop/Udacity-R/R-Project/")
```

```{r}
#Confirming the path of the root folder
getwd()
```

```{r}
#Reading in the CSV data file 
WineQ <- read.csv("wineQualityReds.csv")
```

Checking first few rows of the dataset
```{r}
# Showcases first n rows of dataset
head(WineQ)
```

The data file summary

```{r}
# Provides summary statistics like mean, median for a given attribute in the dataframe
summary(WineQ)
```
**Observation: **
Since mean for none of the variables are NA hence there are no missing values in the dataset.

# Univariate Plot Section

In the univariate analysis we will try to evaluate individual attributes in our dataframe and do some initial data exploration by individual attribute.

Plotting the pH variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = pH), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for the pH variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed.


Transforming the data to make a normal distribution

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(pH)), data = WineQ) + geom_density()
```

Summary before the transformation 

```{r}
# Checking summary before the log transformation
summary(WineQ$pH)
```

Summary after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$pH))
```

**Observation: **
From the above visual and summary statistics we can infer that the distribution is now normally distributed. After the log transformation the scale on the x axis has reduced from 2.7 to 4.0 into 1.0 to 1.3

Plotting fixed.acidity variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = fixed.acidity, fill = I('blue')), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for the fixed.acidity variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Transforming the data into a normal distribution

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
fixed_acidity <-
  ggplot(aes(x = log(fixed.acidity), fill = I('blue')), data = WineQ) +
  geom_density()
fixed_acidity
```

Summary before the transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$fixed.acidity)
```

Summary after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$fixed.acidity))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is now normally distributed. After the log transformation the scale on the x axis has reduced from 4.6 to 15.9 into 1.5 to 2.7

Plotting the volatile.acidity variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = volatile.acidity), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for the volatile.acidity variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Transforming the data into a normal distribution

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(volatile.acidity)) , data = WineQ) + geom_density()
```

Summary before the transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$volatile.acidity)
```

Summary after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$volatile.acidity))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is now normally distributed. After the log transformation the scale on the x axis has reduced from 0.12 to 1.5 into -2.1 to 0.4

Plotting the citric.acid variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = citric.acid, fill = I("red")), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for the citric.acid variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Data transformation 

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
citric_acid <-
  ggplot(aes(x = log(citric.acid), fill = I("red")), data = WineQ) + geom_density()
citric_acid
```

Summarizing the data before transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$citric.acid)
```

Summarizing the data after transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$citric.acid))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is still not normally distributed. After the log transformation the scale on the x axis has reduced from 0.0 to 1 into -Inf to 0. Some alternate transformation procedures need to be followed in order to convert this attribute into a normal distribution. 

Plotting the residual.sugar variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = residual.sugar), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for the residual.sugar variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Transforming the data to a normal distribution

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(residual.sugar)), data = WineQ) + geom_density()
```

Summarizing the data before data transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$residual.sugar)
```

Summarizing the data after data transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$residual.sugar))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is still not normally distributed. After the log transformation the scale on the x axis has reduced from 0.9 to 15.5 into -0.1 to 2.7 Some alternate transformation procedures need to be followed in order to convert this attribute into a normal distribution. 

Plotting the chloride variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = chlorides), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for the chlorides variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Transforming to a normal distribution

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(chlorides)), data = WineQ) + geom_density()
```

Summarizng before the transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$chlorides)
```

Summarizing after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$chlorides))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is now normally distributed. After the log transformation the scale on the x axis has reduced from 0 to 0.6 into -4.4 to -0.4 

Plotting the free.sulfur.dioxide variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = free.sulfur.dioxide), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for free.sulfur.dioxide variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Plotting after the transformation to a normal distribution

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(free.sulfur.dioxide)), data = WineQ) + geom_density()
```

Summarizing before the transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$free.sulfur.dioxide)
```

Summarizing after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$free.sulfur.dioxide))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is still not normally distributed. After the log transformation the scale on the x axis has reduced from 1 to 72 into 0 to 4.2 Some alternate transformation procedures need to be followed in order to convert this attribute into a normal distribution. 

Plotting total.sulfur.dioxide variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = total.sulfur.dioxide), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for total.sulfur.dioxide variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. The bar chart infact helps us to conclude that the distribution is right skewed.

Plotting total.sulfur.dioxide after the data transformation

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(total.sulfur.dioxide)), data = WineQ) + geom_density()
```

Summarizing the data before transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$total.sulfur.dioxide)
```

Summarizing after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$total.sulfur.dioxide))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is normally distributed. After the log transformation the scale on the x axis has reduced from 6 to 289 into 1.7 to 5.6  

Plotting density variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = density), data = WineQ) + 
  geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for density variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed.

Plotting Density after the data transformation

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
ggplot(aes(x = log(density)), data = WineQ) + geom_density()
```

Summarizing density before transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$density)
```

Summarizing after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$density))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is normally distributed. After the log transformation the scale on the x axis has reduced from 0.9 to 1 into -0.0 to 0.0  

Plotting sulphates variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = sulphates, fill = I('Yellow')), data = WineQ) + geom_bar()
```

**Observation: **
The above bar chart visualization shows distribution for sulphates variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. Infact the distribution is right skewed.

Plotting sulphates variable after the transformation

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
sulphates_viz <-
  ggplot(aes(x = log(sulphates), fill = I('Yellow')), data = WineQ) + geom_density()
sulphates_viz
```

Summarizing before transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$sulphates)
```

Summarizing after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$sulphates))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is still not normally distributed. After the log transformation the scale on the x axis has reduced from 0.3 to 2 into -1.1 to 0.6 Some alternate transformation procedures need to be followed in order to convert this attribute into a normal distribution. 

Plotting alcohol variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
ggplot(aes(x = alcohol, color = I('orange')), data = WineQ) + 
  geom_bar()

```

**Observation: **
The above bar chart visualization shows distribution for alcohol variable. Since we observe break points in the distribution hence we can infer that it is not uniformly distributed across all values. Also we can observe that the data is not normally distributed. Infact the distribution is right skewed.

Plotting alcohol variable after the transformation

```{r}
# We will plot a density distribution visual to observe the distribution after log transformation
library("ggplot2")
alcohol_viz <-
  ggplot(aes(x = log(alcohol), fill = I('orange')), data = WineQ) + 
  geom_density()
alcohol_viz
```

Summarizing alcohol variable before the transformation

```{r}
# Checking summary before the log transformation
summary(WineQ$alcohol)
```

Summarizing alcohol variable after the transformation

```{r}
# Checking summary after the log transformation
summary(log(WineQ$alcohol))
```
**Observation: **
From the above visual and summary statistics we can infer that the distribution is still not normally distributed. After the log transformation the scale on the x axis has reduced from 8.4 to 14.9 into 2.1 to 2.7 Some alternate transformation procedures need to be followed in order to convert this attribute into a normal distribution. 

Plotting quality variable

```{r}
# We will plot a bar chart visual to observe the distribution 
library("ggplot2")
quality_viz <-
  ggplot(aes(quality, fill = I("pink")), data = WineQ) + 
  geom_bar()
quality_viz
```

**Observation: **
The above bar chart visualization shows distribution for quality variable. Since we observe no break points in the distribution hence we can infer that it is uniformly distributed across all values. Since it is an ordinal data we won't perform any log transformation for the quality attribute. 

Summarizing the quality variable before transformation

```{r}
# Checking summary for the quality attribute
summary(WineQ$quality)
```

**Observation: **
From the above visual and summary statistics we can infer that the distribution is still not normally distributed. After the log transformation the scale on the x axis has reduced from 3 to 8 into 1 to 2 Some alternate transformation procedures need to be followed in order to convert this attribute into a normal distribution. 

# Bivariate Plots Section

In the bivariate analysis we will try to evaluate the relationship among different attributes in our dataframe and do some data exploration by comparing two attributes.

# Bivariate Analysis

Correlation Matrix

```{r}
# Creation of Correlation Matrix between numeric variables in the dataframe
WineQ <- WineQ[,-1]
Cor_WineQ <- cor(WineQ,use="pairwise.complete.obs")
round(Cor_WineQ, 2)
```

Correlation Plot 

```{r}
# Correlation plot showcasing correlation coefficient and the general data distribution
library(psych)
pairs.panels(WineQ)
```

```{r}
# Correlation plot showcasing correlation coefficient 
library(corrplot)
x <- cor(WineQ)
corrplot(x, type = "upper", order = "hclust", t1.col = "black") 
```

**Observation: **
From the above visualization we can infer that the quality attribute is correlated with alcohol, sulphates and citric.acid. Among these 3 chemical properties quality has correlation coefficient highest with alcohol at 0.48 and reduces with sulphates at 0.25 and citric.acid 0.23. If we observe the correlation coefficients it still does not show a strong relationship (Strong positive correlation will usually have coefficients of 0.5 and above) but in the analysis among the available attributes these 3 variables are correlated with quality of red wine to some extent. Therefore we should try to reduce our scope of analysis and try to explore more with these 3 attributes as far as quality of red wine is concerned. 

Anova Analysis:
Anova is going to compare means of our attribute in consideration across the wines and check if differences are statistically significant compared to the quality of a wine. 

Anova Analysis between alcohol and quality:
Here are the null and alternative hypothesis,
Null Hypothesis (H0): There is no relationship between alcohol and quality of a wine
Alternate Hypothesis (H1): There is a relationship between alcohol and quality of wine

```{r}
# Scatterplot visualization to compare quality vs alcohol
ggplot(aes(x = log(alcohol), y = quality), data = WineQ) +
  geom_jitter(alpha = 0.4) +
  stat_smooth(method = 'lm')
```

**Observation: ** To account for overplotting we have used geom_jitter against geom_point with a transparency set to 0.0000000005 to observe the distribution a bit clearly. 

```{r}
# Using combination of density and scatterplot visualization to compare the relationship
library(gridExtra)
library(grid)
corPlot <-
ggplot(aes(x = log(alcohol), y = quality), data = WineQ) +
geom_jitter(alpha = 0.4) +
stat_smooth(method = 'lm')
grid.arrange(alcohol_viz,
grid.rect(gp = gpar(col = "white")),
corPlot,
quality_viz + coord_flip(),
ncol = 2)
```

```{r}
# Summary statistics after anova analysis
a <- anova(lm(log(alcohol) ~ quality, data = WineQ))
a
```
**Observation: **
As the p value is less than 0.05 we reject the null hypothesis and therefore there is a relationship betweeen quality of wine and alcohol.

Anova Analysis between sulphates and quality:
Here are the null and alternative hypothesis,
Null Hypothesis (H0): There is no relationship between sulphates and quality of a wine
Alternate Hypothesis (H1): There is a relationship between sulphates and quality of wine

```{r}
# Scatterplot visualization to compare quality vs sulphates
ggplot(aes(x = log(sulphates), y = quality), data = WineQ) + 
  geom_jitter(alpha = 0.4) +
  stat_smooth(method = 'lm')
```

**Observation: ** To account for overplotting we have used geom_jitter against geom_point with a transparency set to 0.0000000005 to observe the distribution a bit clearly. 

```{r}
# Using combination of density and scatterplot visualization to compare the relationship
library(gridExtra)
library(grid)
corPlot <-
ggplot(aes(x = log(sulphates), y = quality), data = WineQ) + 
geom_jitter(alpha = 0.4) +
stat_smooth(method = 'lm')
grid.arrange(sulphates_viz,
grid.rect(gp = gpar(col = "white")),
corPlot,
quality_viz + coord_flip(),
ncol = 2)
```

```{r}
# Summary statistics after anova analysis
anova_sul_qual <- anova(lm(log(sulphates) ~ quality, data = WineQ))
anova_sul_qual
```
**Observation: **
As p<0.05 hence we reject null hypothesis and there is a relationship between sulphates and quality of a wine

Anova Analysis between citric.acid and quality:
Here are the null and alternative hypothesis,
Null Hypothesis (H0): There is no relationship between citric.acid and quality of a wine
Alternate Hypothesis (H1): There is a relationship between citric.acid and quality of wine

```{r}
# Scatterplot visualization to compare quality vs citric.acid
ggplot(aes(x = log(citric.acid), y = quality), data = WineQ) + 
  geom_jitter(alpha = 0.4) +
  stat_smooth(method = 'lm')
```

**Observation: ** To account for overplotting we have used geom_jitter against geom_point with a transparency set to 0.0000000005 to observe the distribution a bit clearly. 

```{r}
# Using combination of density and scatterplot visualization to compare the relationship
library(gridExtra)
library(grid)
corPlot <-
ggplot(aes(x = log(citric.acid), y = quality), data = WineQ) + 
geom_jitter(alpha = 0.4) +
stat_smooth(method = 'lm')
grid.arrange(citric_acid,
grid.rect(gp = gpar(col = "white")),
corPlot,
quality_viz + coord_flip(),
ncol = 2)
```

```{r}
# Summary statistics after anova analysis
anova_citric_quality <- anova(lm(citric.acid ~ quality, data = WineQ))
anova_citric_quality
```
**Observation: **
As p<0.05 we reject the null hypothesis and accept the alternate hypothesis. Therefore, there is a relationship between citric.acid and quality of a wine

# Multivariate Plots Section

In the multivariate analysis we will try to evaluate the relationship among different attributes in our dataframe and do some data exploration by comparing multiple attributes (More than 2 attributes). Since we have reduced our scope to 3 attributes : alcohol, sulphates and citric.acid we will try to observe what relationship these variables share with eachother along with the quality attribute in our dataframe.

# Multivariate Analysis

Comparing Citric.Acid vs Sulphates By Quality

```{r}
# Using Scatterplot visual to compare the attributes
ggplot(aes(x = sulphates, y = citric.acid), data = WineQ) +
  geom_jitter(aes(color = quality), alpha = 0.4) +
  geom_smooth(method = lm)
```

**Observation: **
In the above scatterplot visualization we can infer that citric.acid has a strong linear positive relationship with sulphates attribute. We have used lm function to fit linear models. The Linear regression line is plotted at 95% confidence interval.

Comparing Alcohol vs Citric.Acid By Quality

```{r}
# Using Scatterplot visual to compare the attributes
ggplot(aes(x = citric.acid, y = alcohol), data = WineQ) +
  geom_jitter(aes(color = quality), alpha = 0.4) +
  geom_smooth(method = lm)
```

**Observation: ** 
In the above scatterplot visualization we can infer that alcohol has a linear positive relationship with citric.acid attribute but we don't observe a strong relationship. We have used lm function to fit linear models. The Linear regression line is plotted at 95% confidence interval.

Comparing Sulphates vs Alcohol By Quality

```{r}
# Using Scatterplot visual to compare the attributes
ggplot(aes(x = alcohol, y = sulphates), data = WineQ) +
  geom_jitter(aes(color = quality), alpha = 0.4) +
  geom_smooth(method = lm)
```

**Observation: **
In the above scatterplot visualization we can infer that sulphates has a linear positive relationship with alcohol attribute but we don't observe a strong relationship. We have used lm function to fit linear models. The Linear regression line is plotted at 95% confidence interval.

Comparing Citric.acid vs Sulphates vs Alcohol By Quality

```{r}
# Using Scatterplot visual to compare the attributes
ggplot(aes(x = sulphates, y = citric.acid), data = WineQ) +
  geom_jitter(aes(
  color = quality,
  size = alcohol),
  alpha = 0.4
  ) +
  geom_smooth(method = lm) 
```

**Observation: **
In the above scatterplot visualization we can infer that sulphates has a strong linear positive relationship with citric.acid attribute. We have used lm function to fit linear models. The Linear regression line is plotted at 95% confidence interval. We can conclude that major red wines with high alcohol and high quality content are in sulphates between 0.5-1 and citric.acid 0-0.75 range.

# Final Plots and Summary
In this section I have shortlisted three visualizations of the highest significance which helped me analyze the chemical properties which affect the quality of red wine. 

### Plot One

```{r}
# Correlation plot showcasing correlation coefficient 
library(corrplot)
title <- "Correlation Plot"
x <- cor(WineQ)
corrplot(x, type = "upper", order = "hclust", t1.col = "black", title = title, mar=c(0,0,1,1)) 
```

### Description One

Improvements made in the visualization,
1. A title was added to the visualization and title alignment was appropriately set.
Reason for selection of the visualization,
1. The correlation plot helped to reduce the scope of analysis by providing us the 3 variables which had the strongest relationship with the quality attribute. 

### Plot Two

```{r}
# Using Scatterplot visual to compare the attributes
ggplot(aes(x = sulphates, y = citric.acid), data = WineQ) +
  geom_jitter(aes(color = quality), alpha = 0.4) +
  geom_smooth(method = lm) +
  ggtitle(" Comparing citric.acid vs sulphates by quality ") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("sulphates (g/dm3)") +
  ylab("citric acid (g/dm^3)")
```

### Description Two

Improvements made in the visualization,
1. A title was added to the visualization and title alignment was appropriately set.
2. Appropriate axis labels were included. (Includes both x and y axis)
Reason for selection of the visualization,
1. After the correlation plot i wanted to have an understanding on what were the relationships shared among the three variables. This visualization helped me to understand that a strong linear positive relationship is associated between citric.acid and sulphates. Therefore if we aggregate other attributes on top of this visualization it will be easier to get an understanding on what chemical factors and to what extent affect the quality of red wine. 

### Plot Three

```{r}
# Using Scatterplot visual to compare the attributes
ggplot(aes(x = sulphates, y = citric.acid), data = WineQ) +
  geom_jitter(aes(
  color = quality,
  size = alcohol),
  alpha = 0.4
  ) +
  geom_smooth(method = lm) +
  ggtitle(" Comparing citric.acid vs sulphates by quality & alcohol") +
  theme(plot.title = element_text(hjust = 0.5)) +
  xlab("sulphates (g/dm3)") +
  ylab("citric acid (g/dm^3)")
```

### Description Three

Improvements made in the visualization,
1. A title was added to the visualization and title alignment was appropriately set.
2. Appropriate axis labels were included. (Includes both x and y axis)
Reason for selection of the visualization,
1. After understanding that citric acid and sulphates have a strong positive linear relationship i wanted to include other attributes of our interest. In this case as we knew alcohol is a chemical property of interest we included it as a size. Which helped us to conclude that major red wines with high alcohol and high quality content are in sulphates between 0.5-1 (g/dm3) and citric.acid 0-0.75 (g/dm^3) range.

# Reflection

What went well?
Since the dataset was clean and without missing values it was easy to start the data exploration peice without any significant cleaning.
What was surprising?
Before starting the data exploration to my limited understanding on wines i thought residual.sugar along with alcohol could be the attributes of our interest as far as quality of wine is considered but while evaluating the data i got clarified that residual.sugar does not have a strong correlation with quality and second aspect of surprise was that alcohol had the strongest correlation coefficient when it comes to quality of red wine.
Future scope:
There could be machine learning algorithms applied to the dataset for example:
Outlier detection algorithms could be used to detect the few excellent or poor wines. We could also perform some classification algorithm to classify excellent and poor wines. 

# References

http://rcompanion.org/handbook/I_12.html
http://ggplot2.tidyverse.org/reference/geom_density.html
http://www.analyticsforfun.com/2014/06/performing-anova-test-in-r-results-and.html
http://www.imachordata.com/extra-extra-get-your-gridextra/
http://www.sthda.com/english/wiki/correlation-matrix-a-quick-start-guide-to-analyze-format-and-visualize-a-correlation-matrix-using-r-software
https://stackoverflow.com/questions/5446426/calculate-correlation-for-more-than-two-variables





